---
title: "Importing and Graphing Network Data"
author: "Dr. Tracy Sweet"
date: 
output: html_document
---


## Importing from a Previously Saved R Workspace
We'll import some network data. 

```{r}
load("WorkshopData.Rdata")
```

# Plotting Network Data in R
Now that we have our data into R, the easiest way to plot our data in R is to use the igraph package.  We also have to create igraph objects.


## Making an Igraph Object
```{r}
library(igraph)
teachergraph=graph.adjacency(as.matrix(Ela9Network), mode="directed")
authorgraph=graph.adjacency(as.matrix(authornet), mode="undirected")
```


## Using the Igraph Default Plot

Sometimes the default ploting function makes perfectly adequate plots and other times, the plots are terrible.  You be the judge.
```{r}
plot(teachergraph)
plot(authorgraph)

```


## Let's redo the teacher plot with some changes
```{r}
plot(teachergraph, vertex.size=15, vertex.color="orange", vertex.frame.color="NA", edge.width=1, vextex.label=1:13)

plot(teachergraph, vertex.size=8, vertex.label=NA, vertex.color="darkgreen", vertex.frame.color="NA", edge.arrow.size=0.5,  edge.curved=0.2)

## What is I wanted to save the node positions?
coords=layout_with_fr(teachergraph) ## this is the most common layout
plot(teachergraph, vertex.size=15, vertex.frame.color=NA,  vertex.color="darkgreen", vertex.label.color="white", edge.arrow.size=0.5, layout=coords)

## To remove arrows
plot(teachergraph, vertex.size=15, vertex.frame.color=NA, vertex.label=1:24, vertex.color="darkgreen", vertex.label.color="white", vertex.label=1:24, edge.arrow.size=0.5, layout=coords, edge.arrow.mode=0)

## To color the nodes based on grade taught
mycolors=rainbow(8)
plot(teachergraph, vertex.size=15, vertex.frame.color=NA, vertex.label=1:24, vertex.color=mycolors[Teacher9$grade], vertex.label.color="white", vertex.label=1:24, edge.arrow.size=0.5, layout=coords, edge.arrow.mode=0)
## We can then save these plots as pdfs!
```



## Redoing the CoAuthorship Network with some changes
```{r}
plot(authorgraph, vertex.size=12, vertex.frame.color=NA, vertex.color="gold", vertex.label=NA, edge.arrow.mode=0)

coords=layout_nicely(authorgraph) 
plot(authorgraph, vertex.size=12, vertex.frame.color=NA, vertex.color="gold", vertex.label=NA, edge.arrow.mode=0, layout=coords)
## Sometimes it doesn't always work and just rerunning the code will get you a nicer looking plot.

## Leaders are bigger
plot(authorgraph, vertex.size=authordf$YrsPostPhD, vertex.frame.color=NA,
    vertex.color="gold", vertex.label=NA, edge.arrow.mode=0)

## Faculty are colored differently
mycolors=c("gold", "red")
faculty=authordf$CurrentFaculty+1
plot(authorgraph, vertex.size=15, vertex.frame.color=NA,
    vertex.color=mycolors[faculty], vertex.label=NA, edge.arrow.mode=0)

```

## Descriptive Statistics 

### Describing the Network
Density, reciprocity and transitivity
```{r}
teachergraph=graph.adjacency(Ela9Network)
edge_density(teachergraph)
edge_density(authorgraph)
reciprocity(teachergraph)

## Remember this transitivity
## function forces all ties to bidirectional!
transitivity(teachergraph)
transitivity(authorgraph)

```


### Describing the Nodes


To calculate degree, use the argument "mode" to determine whether you want to calculate in-degree, out-degree or degree (sum of in and out).  Recall that in and out degree are for directed networks only.What you get is the in-, out-, or total degree for each node.  Note that if you don't select a mode, you get the sum by default.
```{r}
degree(teachergraph, mode="in")
degree(teachergraph, mode="out")
degree(authorgraph)
```

Closeness and betweenness are also directional since it's based on the distance between pairs of nodes.  
```{r}
## Recall that closeness and betweenness don't work well for disconnected graphs
closeness(teachergraph, mode="out") 
betweenness(teachergraph, directed=TRUE)
betweenness(authorgraph, directed=FALSE)

```


### Edge Statistics
There are a number of distance functions in igraph.
See http://igraph.org/r/doc/distances.html for more information.
To calculate the geodesic distance (shortest path) between every pair of nodes.
```{r}
## We typically use out for geodesic distance for directed networks 
distances(teachergraph, mode="out")
## Note you can choose "in" or "all"
## the default option is "all" which is 
## appropriate for undirected networks
```



## Community Detection
```{r}
library(igraph)
teachergraph=graph.adjacency(Ela9Network)
```

### Random Walk Community Detection
```{r}
wtc=cluster_walktrap(teachergraph, steps=4)

## membership returns a membership id for each node
wtc.membership=membership(wtc)
wtc.membership

## communities returns a list of node names in each block
communities(wtc)

plot(teachergraph, vertex.color=wtc.membership, vertex.size=12)

## Let's try this again with a larger number of steps
wtc=cluster_walktrap(teachergraph, steps=6)
wtc.membership2=membership(wtc)
plot(teachergraph, vertex.color=wtc.membership2, vertex.size=12)
```

### Edge Betweenness
```{r}
ebc=cluster_edge_betweenness(teachergraph)
## Notice that the membership for the highest modularity is returned ##
membership(ebc)
ebc.membership=membership(ebc)
plot(teachergraph, vertex.color=ebc.membership, vertex.size=12)
```

### Fast Greedy
We'll need to change the network slightly so we can work with this algorithm.  Let's change the network.
```{r}
teachergraph2=graph.adjacency(Ela9Network, mode="undirected")
fgc=cluster_fast_greedy(teachergraph2)
fgc.membership=membership(fgc)
plot(teachergraph2, vertex.color=fgc.membership, vertex.size=12)

```


